Implications of afforestation policies for global markets

### 2024-02-27 paper.Rmd
Update paper.Rmd

Afforestation efforts are also thought to not provide equal benefits in every region. For example, afforestation in boreal zones has been shown to provide the smallest net benefits, as the carbon sequestrations benefits of conifer forests are mitigated by the reduction in albedo in winter [@mykleby2017; @li2015]. Coniferous forests have also been shown to increase soil organic carbon less than broadleaf forests [@guo2021]. Conversely, tropical forests have been linked with a strong positive result from afforestation, as the carbon sequestration benefit is complemented by a cooling impact due to changes in both albedo and evapotranspiration, and with temperate forests, the impact of afforestation is also shown to be positive, albeit less strongly so due to a winter warming effect [@li2015].

South America and Sub-Saharan Africa are thus two regions that show much potential for afforestation. It is estimated that these two regions hold at least 50% of the potential global gain while regions such as Northern Africa and the Middle East show little promise as forest growth rates are quite low [@doelman2020]. Thus, it is promising that the African Forest Landscape Restoration Initiative (AFR100), which aims to afforest 100 million hectares by 2030, has roughly 70% of its commitments coming from Sub-Saharan nations [@auda2023]. While many of these nations may be well-suited in biophysical terms for afforestation, there is concern about how long-lasting any investment may be due to proximate zones of political instability. There are other nations that have a big potential for gains from afforestation, but they also face challenges. India, for example, has a great potential to increase its carbon sequestration through afforestation, but its massive population and food security worries limit its ability to consider such actions. The United Nations suggests that India's population will grow slightly less fast than the world average between 2023 and 2033 (+8.5%), but due to its sizable base, that will equate to an increase in population of 120.5 million. Thus, its NDC does not entertain the idea of agricultural land being re-purposed and instead primarily considers targeting wastelands that may not be well suited for afforestation [@amjath2019]. Moreover, as land-use decisions in one country also tend to have impacts that spillover into many others [@meyfroidt2010; @lambin2011], analysis needs to be conducted at a global scale.

@@

The rationale for joining these two models is that, by passing outputs back and forth between the models until an dual-state equilibrium is achieved, any limitations of either model should be minimized and more robust results should, therefore, be attained. Establishing a connection between the  GTAP model and the GTM model required a two-step process. First, we had to identify the set of variables that are common to both models and their closures. From those, we selected which variables would be determined endogenously by one model and held as exogenous in the other model. In our case, we decided to model the return on land using the GTAP model and feed those results into the GTM model as exogenous. The GTM model then endogenously determined the amount of land available for crop and livestock production, and we fed that back into the GTAP model as an exogenous input. In the second step, we established the connection between the models on the software level. This step required that the GTAP model, which is written in GEMPACK, be able to handle inputs from the GTM model, which is written in GAMS, and submit the outputs back to the GTM model. As GEMPACK and GAMS were developed long before software integration became important or even possible, we settled on the use of R [@R2021] to create the user functions that executed both models, collected their outputs, and automatically created parameter files.

>>>>

It is important to understand where to target afforestation efforts, as they are thought to not provide equal benefits in every region. For example, afforestation in boreal zones has been shown to provide the smallest net benefits, as the carbon sequestrations benefits of conifer forests are mitigated by the reduction in albedo in winter [@mykleby2017; @li2015]. Coniferous forests have also been shown to increase soil organic carbon less than broadleaf forests [@guo2021]. Conversely, tropical forests have been linked with a strong positive result from afforestation, as the carbon sequestration benefit is complemented by a cooling impact due to changes in both albedo and evapotranspiration, and with temperate forests, the impact of afforestation is also shown to be positive, albeit less strongly so due to a winter warming effect [@li2015].

South America and Sub-Saharan Africa are thus two regions that show much potential for afforestation. It is estimated that these two regions hold at least 50% of the potential for global gains while regions such as Northern Africa and the Middle East show little promise as forest growth rates are quite low [@doelman2020]. Thus, it is promising that the African Forest Landscape Restoration Initiative (AFR100), which aims to afforest 100 million hectares by 2030, has roughly 70% of its commitments coming from Sub-Saharan nations [@auda2023]. While many of these nations may be well-suited in biophysical terms for afforestation, there is concern about how long-lasting any investment may be due to proximate zones of political instability. There are other nations that have a big potential for gains from afforestation, but they also face challenges. India, for example, has a great potential to increase its carbon sequestration through afforestation, but its massive population and food security worries limit its ability to consider such actions. The United Nations suggests that India's population will grow slightly less fast than the world average between 2023 and 2033 (+8.5%), but due to its sizable base, that will equate to an increase in population of 120.5 million. Thus, its NDC does not entertain the idea of agricultural land being re-purposed and instead primarily considers targeting wastelands that may not be well suited for afforestation [@amjath2019]. Moreover, as land-use decisions in one country also tend to have impacts that spillover into many others [@meyfroidt2010; @lambin2011], analysis needs to be conducted at a global scale.

@@

The rationale for joining these two models is that, by passing outputs back and forth between the models until a dual-state equilibrium is achieved, any limitations of either model should be minimized and more robust results should, therefore, be attained. Establishing a connection between the  GTAP model and the GTM model required a two-step process. First, we had to identify the set of variables that are common to both models and their closures. From those, we selected which variables would be determined endogenously by one model and held as exogenous in the other model. In our case, we decided to model the return on land using the GTAP model and feed those results into the GTM model as exogenous. The GTM model then endogenously determined the amount of land available for crop and livestock production, and we fed that back into the GTAP model as an exogenous input. In the second step, we established the connection between the models on the software level. This step required that the GTAP model, which is written in GEMPACK, be able to handle inputs from the GTM model, which is written in GAMS, and submit the outputs back to the GTM model. As GEMPACK and GAMS were developed long before software integration became important or even possible, we settled on the use of R [@R2021] to create the user functions that executed both models, collected their outputs, and automatically created parameter files.
